
## "Chapter3"

```{r}
library(GGally)
library(ggplot2)
library(tidyr); 
```

```{r}
url<-"http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt"
```

```{r}
colnames(alc)
str(alc)
dim(alc)
```

#I have chosen internet, traveltime, age and freetime I predict that internet use increases alcohol consumpiton. I predict that traveltime has no correlation. I predict that age is related to alchohol consumption as well as the amount of ones freetime.

```{r}
summary(alc$studytime)
summary(alc$traveltime)
summary(alc$age)
summary(alc$freetime)
```


```{r}
alc_data <- select(alc,c("high_use","studytime","traveltime","age","freetime"))
gather(alc_data) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

```{r}
p<-ggpairs(alc_data)
p
```


```{r}
m <- glm(high_use ~ high_use + studytime + traveltime + age + freetime, data = alc_data, family = "binomial")
summary(m)
coef(m)
```


```{r}
OR <- coef(m) %>% exp
CI <- confint(m) %>% exp
cbind(OR, CI)
```

# predict() the probability of high_use
```{r}
probabilities <- predict(m, type = "response")
```

# add the predicted probabilities to 'alc'
```{r}
alc_data <- mutate(alc_data, probability = probabilities)
```

# use the probabilities to make a prediction of high_use
```{r}
alc_data <- mutate(alc_data, prediction = probability > 0.5)
```

# see the last ten original classes, predicted probabilities, and class predictions
```{r}
select(alc_data, high_use, studytime, traveltime, age, freetime) %>% tail(10)
```

```{r}
table(high_use = alc_data$high_use, prediction = alc_data$prediction)
```

loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}
loss_func(class = alc_data$high_use, prob = alc_data$probability)

library.tidyr

library(boot)
cv <- cv.glm(data=alc_data, cost=loss_func, glmfit=m, K=10)
cv$delta[1]

datacamp had [1] 0.2539267 as cv value.
