---
title: "Chapter5"
output: html_document
---
```{r}
human = read.table('/Users/eric/Documents/GitHub/IODS-project/data/human.txt',sep='\t',header = TRUE)

str(human)
dim(human)
```
The human data is from the UNITED NATIONS DEVELOPMENT PROGRAMME Human Development Reports 2016
"The HDI(human development index) was created to show that people and their abilities should be the ultimate criteria for assessing the development of a country, not economic growth alone. The HDI has the potential to also be used to question national policy choices, asking how two countries with the same level of GNI per capita can end up with different human development outcomes."
# Accessed libraries
```{r}
library(MASS)
library(corrplot)
library(tidyr)
library(GGally)
library(ggplot2)
```
#Visualized human
```{r}
ggpairs(human)

cor_matrix=cor(human)
corrplot(cor_matrix,type='upper')
```
From the scatter plot, it is seen that Edu.Exp and Life.Exp  have a strong postivie correlation (0.789) as well as Mat.Mor and Ado.Birth. While Life.Exp and Mat.Mor have a strong negative correlation(-0.857).

In the corrplot Positive correlations are displayed in blue and negative correlations in red color. Color intensity and the size of the circle are proportional to the correlation coefficients. In the right side of the correlogram, the legend color shows the correlation coefficients and the corresponding colors. The corrplot shows the correlations between variables of the Boston dataset. The highest postive correlation can be seen between Life.Exp and Edu.Exp and the highest negative correlatiuon is seen between Life.Exp and Math.Mor
#Perform Principal Component Analysis (PCA) non-standardized variables

```{r}
pca_human <- prcomp(human)
s <- summary(pca_human)
s
pca_pr=round(100*s$importance[2,],digits=1)
pc_lab=paste0(names(pca_pr),'(',pca_pr,'%)')
biplot(pca_human,choices=1:2,cex=c(0.8,1),xlab=pc_lab[1],ylab=pc_lab[2])
```
With the non-standardized variables, it is difficult to read the bi-plot. The scales are extreme (-200000 and -0.4).
PC1 is 100% and PC2 is 0%
#Perform Principal Component Analysis (PCA) on standardized variables
```{r}
human_std <- scale(human)
pca_human=prcomp(human_std)
s=summary(pca_human)
s
pca_pr=round(100*s$importance[2,],digits=1)
pc_lab=paste0(names(pca_pr),'(',pca_pr,'%)')
biplot(pca_human,choices = 1:2,cex=c(0.8,1),xlab=pc_lab[1],ylab=pc_lab[2])
```
The biplots between standardized and nonstandardized variables have different results because the biplot with the non-standardized variable had extreme variance while the biplot wtih standardized variables, all of the variables have similiar variance. It is easy to see that between the two biplots, there is a big difference between the distribution of the data. 

PC1 is 53.6% and PC2 is 16.2%.


It is clearly show in the biplot that Mat.Mor and Ado.Birth have a positive relationship between each other. It is also show that Edu.Exp,Edu2.FM,GNI and Life.Exp have a positive relationship between other.

While Palri.f and Labo.fm have have a negative relationship.


```{r}
library(FactoMineR)
library(ggplot2) 
library(dplyr) 
library(tidyr)
```


```{r}
data('tea')
str(tea)
dim(tea)
```

The data set tea contains 300 observations and 36 variables, many of them are factors.

Create and visualized tea_time 
```{r}
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- dplyr::select(tea, one_of(keep_columns))
summary(tea_time)
str(tea_time)

gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```
Created tea_time from tea by selecting the columns "Tea", "How", "how", "sugar", "where", "lunch". 
In Lunch, Not Lunch has a higher count then lunch. 
In sugar, No sugar's count is close to sugar. 
In where, chain store has the highest count.
In How, alone has the highest count.
In tea, earl grey has the highest count.
In how, tea bag has the highest count.

Perform MCA on tea_time
```{r}
mca=MCA(tea_time,graph=FALSE)
summary(mca)
plot(mca,invisible=c('ind'),habillage='quali')
```
From the summary; tea bag, earl grey, and lemon are most likely Dim3. 
lone, green, milk, black tea, tea bag+unpackaged, unpackaged, and other are most likely Dim1. It seems that lunch, How, Tea, and sugar are most likely in Dim3. where and how are most likely Dim.1 while where could also be Dim2.

MCA is generally used to analyse a data set from survey. The goal is to identify a group of individuals with similar profile in their answers to the questions and the associations between variable categories
From the MCA graph, it is clear that tea bag + unpackaged is the most associated with Dim1 while other is the most associated with Dim2.
